Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-14 prop_letter.tex
prop_letter.tex \section{Gleason} Updating...

The QIVPM~$\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$ is well-behavior
in a sense that the collection of states consistent it get closer
and closer to $\rho$ as the intervals get narrower and narrower,
eventually collapsing to $\rho$ when the intervals become infinitely
precise real numbers. To state this result, we need some more technical
notions. We first define a notion of distance based on the following
norm of a matrix~$A$ as $\left\Vert A\right\Vert =\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{A}{\psi}\right|$~\citep{544199}.
Then, for any state~$\rho$ satisfying $0<\mu_{\rho}^{\mathrm{B}}\left(P\right)<1$
for any $P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $,
any other state~$\rho'$ consistent with $\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$
must close to $\rho$ in a sense that $\left\Vert \rho-\rho'\right\Vert \le1/\left(12n\right)$.
Reversely, if a state~$\rho'$ is close to $\rho$ in a sense that
$\left\Vert \rho-\rho'\right\Vert \le1/n$, then $\rho'$ is consistent
with $\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$. In other
words, for the states consistent with this QIVPM, $\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$
will become closer and closer to $\rho'$ as $n$ goes to infinity.
In the limiting case when the intervals are infinitely precise, i.e.,
$\set{\left[x,x\right]}{x\in\left[0,1\right]}$, any QIVPM can be
recast as a real-valued one and will have a unique state consistent
with it on every projector.

\yutsung{Clean up until here.}

In the classical setting, measurements done with limited resources
and poor precision may not uniquely identify the true state of a system.
However, there is always at least one state that is consistent with
the imprecise measurements.

\begin{thm}[Shapley~\cite{Shapley1971,GilboaSchmeidler1994,Grabisch2016}]
  For a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$, if a subspace
  of events~$\events'\subseteq\events$ commutes, then
  $\coreBorn\left(\bar{\mu},\events'\right)\ne\emptyset$.
\end{thm}

In the idealized situation of real-valued quantum probability
measures, Gleason's theorem asserts that every measure is determined
by a unique state.

\begin{thm}[Gleason~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}]\label{cor:Gleason's}In
a Hilbert space $\Hilb$ of dimension $d\geq3$, given a quantum probability
measure~$\mu:\events\rightarrow[0,1]$, there exists a unique 
state~$\rho$ such that $\mu=\mu_{\rho}^{\mathrm{B}}$.
\end{thm}

Thm.~\ref{thm:Shapley} is weaker: it asserts that, for each classical
subspace of events, a QIVPM must be consistent with \emph{some}
state. But the states consistent with different subspaces may be
incomparable, as the following example shows.

Given a three-dimensional Hilbert space $\Hilb$ with an orthonormal
basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $, let
$\mathscr{I}_{1}=\left\{ \necess,\imposs,\unknown\right\} $,
$\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$,
$\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$, and
\begin{subequations}
\begin{eqnarray}
 &  & \begin{aligned}\imposs & =\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})\\
 & =\bar{\mu}(\proj{\ps})=\bar{\mu}(\proj{\ps'})\,,
\end{aligned}
\\
 &  & \begin{aligned}\necess & =\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})\\
 & =\bar{\mu}(\mathbb{1}-\proj{\ps})=\bar{\mu}(\mathbb{1}-\proj{\ps'})\,,
\end{aligned}
\\
 &  & \unknown=\bar{\mu}(P)\qquad\textrm{otherwise.}
\end{eqnarray}
\end{subequations}
It is straightforward to verify that $\bar{\mu}$ is a QIVPM. Then we
will prove $\coreBorn\left(\bar{\mu},\events\right)$ is an empty set
by contradiction. Suppose there is a state
$\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}\in\coreBorn\left(\bar{\mu},\events\right)$,
where $\sum_{j=1}^{N}q_{j}=1$ and $q_{j}>0$. Since
$\mu_{\rho}^{\mathrm{B}}(P)\in\bar{\mu}(P)$ for any~$P$, the
state~$\rho$ must satisfy:
\begin{equation}
\mu_{\rho}^{\mathrm{B}}(\proj{0})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps'})=0
\end{equation}
which implies
$\ip{\phi_{j}}{0}=\ip{\phi_{j}}{\ps}=\ip{\phi_{j}}{\ps'}=0$ for all
$j$. However, the last equation never holds.  Thus, there is no state
consistent with the QIVPM on all projectors~$P\in\events$, so
Gleason's theorem does not hold, at least for this example.

>>>>

The QIVPM~$\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$ is well-behavior
in a sense that the collection of states consistent it get closer
and closer to $\rho$ as the intervals get narrower and narrower,
eventually collapsing to $\rho$ when the intervals become infinitely
precise real numbers. To state this result, we need some more technical
notions. We first define a notion of distance based on the following
norm of a matrix~$A$ as $\left\Vert A\right\Vert =\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{A}{\psi}\right|$~\citep{544199}.
Then, for any state~$\rho$ satisfying $0<\mu_{\rho}^{\mathrm{B}}\left(P\right)<1$
for any $P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $,
any other state~$\rho'$ consistent with $\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$
must close to $\rho$ in a sense that $\left\Vert \rho-\rho'\right\Vert \le1/\left(12n\right)$.
Reversely, if a state~$\rho'$ is close to $\rho$ in a sense that
$\left\Vert \rho-\rho'\right\Vert \le1/n$, then $\rho'$ is consistent
with $\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$. In other
words, for the states consistent with this QIVPM, $\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$
will become closer and closer to $\rho'$ as $n$ goes to infinity.

In the limiting case when the intervals are infinitely precise, i.e.,
$\set{\left[x,x\right]}{x\in\left[0,1\right]}$, even if a QIVPM is
not \emph{a priori} known to be induced by a state, it can be recast
as a real-valued one and will have a unique state consistent with
it on every projector. This fact is called Gleason's theorem.

\begin{thm}[Gleason~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}]\label{cor:Gleason's}In
a Hilbert space $\Hilb$ of dimension $d\geq3$, given a quantum probability
measure~$\mu:\events\rightarrow[0,1]$, there exists a unique state~$\rho$
such that $\mu=\mu_{\rho}^{\mathrm{B}}$. \end{thm}

\noindent However, if the intervals are only with finite-precision,
there are some QIVPMs which cannot be consistent with any state. Given
a three-dimensional Hilbert space $\Hilb$ with an orthonormal basis
$\left\{ \ket{0},\ket{1},\ket{2}\right\} $, let $\mathscr{I}_{1}=\left\{ \necess,\imposs,\unknown\right\} $,
$\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$, $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$,
and \begin{subequations} 
\begin{eqnarray}
 &  & \begin{aligned}\imposs & =\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})\\
 & =\bar{\mu}(\proj{\ps})=\bar{\mu}(\proj{\ps'})\,,
\end{aligned}
\\
 &  & \begin{aligned}\necess & =\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})\\
 & =\bar{\mu}(\mathbb{1}-\proj{\ps})=\bar{\mu}(\mathbb{1}-\proj{\ps'})\,,
\end{aligned}
\\
 &  & \unknown=\bar{\mu}(P)\qquad\textrm{otherwise.}
\end{eqnarray}
\end{subequations} It is straightforward to verify that $\bar{\mu}$
is a QIVPM. Then we will prove $\coreBorn\left(\bar{\mu},\events\right)$
is an empty set by contradiction. Suppose there is a state $\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}\in\coreBorn\left(\bar{\mu},\events\right)$,
where $\sum_{j=1}^{N}q_{j}=1$ and $q_{j}>0$. Since $\mu_{\rho}^{\mathrm{B}}(P)\in\bar{\mu}(P)$
for any~$P$, the state~$\rho$ must satisfy: 
\begin{equation}
\mu_{\rho}^{\mathrm{B}}(\proj{0})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps'})=0
\end{equation}
which implies $\ip{\phi_{j}}{0}=\ip{\phi_{j}}{\ps}=\ip{\phi_{j}}{\ps'}=0$
for all $j$. However, the last equation never holds. Thus, there
is no state consistent with the QIVPM on all projectors~$P\in\events$,
so Gleason's theorem cannot be extended to general QIVPMs because
of this example.
